---
title: "STAT 385 - Homework 6"
author: "Ali Asghar - aasgha2"
date: "Due 11:59 PM, 3/30/2020"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***

## Submit your homework

Make sure you check the following requirements when submitting your homework:

- Put your name - your NetID as the author info at the top of the Rmd file.
- After pushing the latest commit to Git, check if the changes are reflected on [github.com](https://github.com/).
  - DO NOT directly upload to github.com! That is not how we push a commit in this class!

Failure to meet the above requirements will negatively impact your grade.

***

## Before you start

- **For all plots in this homework, you have to use the functions from the ggplot2 package**.
  - If you use base R plotting functions, your answers will NOT be graded.
- The maximum possible points for this homework is 70.
  - Make sure you answer all the questions!

***

## Dataset: diamonds from the ggplot2 package

The ggplot2 package include many interesting datasets. We looked at `mpg` in lecture this week. Now, we will explore the `diamonds` dataset. This data contains the prices and other attributes of almost 54,000 diamonds.

```{r message=FALSE}
library(tidyverse)
str(diamonds)
```

- Run `?diamonds` to know what each variable presents.

### Scatterplot

- There are 4 parts to this exercise: each part worths 5 points.

- **(a)** Graph a scatterplot of `price` vs `carat`.
  - `price` as the $y$-axis and `carat` as the $x$-axis.
  
```{r fig.align='center'}
# add your graph here
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price))
```

- **(b) With the above graph**, adjust the transparency level of ALL the points so we can see clearly where the clusters are:

```{r fig.align='center'}
# add your graph here
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price), alpha = 0.5)
```

- **(c) Now create a new plot**, graph a scatterplot of `price` vs `carat`, and color the points BY the quality of the cut (`cut`)!
  - Don't change the transparency level of the points.

```{r fig.align='center'}
# add your graph here
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price, color = cut))
```

- **(d) What can you conclude based on the above graph?**
  - **ANSWER: A diamond of lower carat value tends to have better quality cuts** 
  
### Boxplot

- There are 3 parts to this exercise: each part worths 5 points.

- **(a)** Graph a boxplot of `carat` vs `cut`.
  - `carat` as the $y$-axis and `cut` as the $x$-axis.
  
```{r fig.align='center'}
# add your graph here
sp <- ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = carat, color = cut))
sp
```

- **(b) With the plot above**, rotate it so `cut` is on the $y$-axis and `carat` is on the $x$-axis.
  - Make sure the plot is still a boxplot.
  
```{r fig.align='center'}
# add your graph here
sp + coord_flip()
```

- **(c) With the plot from (b)**, color the boxplot dark blue.

```{r fig.align='center'}
# add your graph here
sp2 <- ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = carat), color = "dark blue")
sp2 + coord_flip()
```

### Histogram

- There are 3 parts to this exercise: each part worths 5 points.

- **(a)** Graph a histogram of `price`.
  
```{r fig.align='center'}
# add your graph here
ggplot(data = diamonds, aes(x=price)) +
  geom_histogram(binwidth=100)
```

- **(b) With the plot above**, color the bars dark orange.
  
```{r fig.align='center'}
# add your graph here
bp <- ggplot(data = diamonds, aes(x=price)) +
  geom_histogram(binwidth=100, color = "dark orange", fill = "dark orange")
bp
```

- **(c) With the plot from (b)**, change the number of bins to 50.

```{r fig.align='center'}
# add your graph here
ggplot(data = diamonds, aes(x=price)) +
  geom_histogram(binwidth=50, color = "dark orange", fill = "dark orange")
```



***

## Dataset: 2015 Flight Delays and Cancellations Data

In this exercise, we continue to work with the 2015 Flight Delays and Cancellations Data provided by the U.S. Department of Transportation. This is a huge dataset availalbe on [Kaggle](https://www.kaggle.com/usdot/flight-delays#airports.csv). But for us, we will only take a look at flights **flying out** from O'Hare International Airport (ORD) in January, 2015.

### Load data

- I have filtered out the data specific to O'Hare and stored it in `ohare_jan.csv`. This filtered data is available at the URL: https://nkha149.github.io/stat385-sp2020/files/data/ohare_jan.csv.

```{r}
flights <- read_csv(file = "https://nkha149.github.io/stat385-sp2020/files/data/ohare_jan.csv")
```

### Using `dplyr` and `ggplot2` together

- For this exercise, you will need to:
  - Graph a plot that can be used to answer the question.
  - Determine the type of plot that is needed.
  - (You might need to) Manipulate the data to get the specific information that you want.
  
- There are 2 parts to this exercise: each part worths 10 points.

- **(a)** How does the **average** departure delay time differ among days of the week?



```{r}
data1 <- flights %>%
  group_by(DAY_OF_WEEK) %>%
  summarise(ave_delay = mean(DEPARTURE_DELAY, na.rm = TRUE),
            n = n())
ggplot(data = data1) +
  geom_point(mapping = aes(x = DAY_OF_WEEK, y = ave_delay))
```
  
- **(b)** Airlines are grouped into different terminals. There is a belief that certain terminals have better taxing out time. How does **median** taxi out time differ among diferrent airlines?







